Timing and Likelihood of the Origin of Life Derived from Post-Impact Highly Reducing Atmospheres

### 2023-08-07 ms.tex
forgot some bits

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/probabilities_of_impacts.pdf}
  \caption{The probability of an impact-driven origin of life on Earth. The blue solid line is the probability of an impact occurring at least once. The orange dashed line is the probability of an impact without a subsequent ocean-vaporizing collision, which can be interpreted as the chance of an impact that starts life. Probabilities are shown as a function of the minimum impact mass to produce significant prebiotic molecules (e.g. HCN). Two plausible minimum masses are the \citet{Wogan_2023} optimistic and pessimistic scenarios indicated with vertical dotted lines. The shaded region between $2 \times 10^{22}$ and $6 \times 10^{22}$ kg are Moneta-sized impacts because they deliver most all of Earth's mantle HSEs. There is a 92\% chance of a life-starting impact in the \citet{Wogan_2023} optimistic case, and a 41\% chance in the pessimistic case.}
  \label{fig:probabilities_of_impacts}
\end{figure}

@@

It is temping to extrapolate our results beyond the early Earth to exoplanets, but we must do so with caution because planet's orbiting different stars may have bombardment histories unlike the Hadean. Planets in the habitable zone of an M-type star during the stellar main-sequence phase were interior to the habitable zone for several hundred million years during the super luminous pre-main-sequence phase \citep{Luger_2015}. \citet{Lichtenberg_2022} uses N-body simulations to show that, for planets hosted by M-dwarfs, large asteroid impacts are not useful for prebiotic chemistry because big impacts most likely occur within 100 million years of planet formation during the stellar pre-main-sequence when the planet is outside of the habitable zone. The pre-main-sequence phase is not an issue for an impact-induced origin of life on planets orbiting sun-like stars because the phase only lasts 10s Myrs, while large asteroid impacts occur for 100s Myrs \citep{Lichtenberg_2022}. Therefore, our result that a life-starting impact was likely on the early Earth (92\% chance in an optimistic case) might be best extrapolated to habitable zone exoplanets orbiting sun-like stars.

@@

Our simulations of Earth's impact history do not always result in an origin of life. There are some impact histories where a collision of sufficient mass to produce prebiotic molecules does not occur, or alternatively, does occur but primitive life is subsequently destroyed by an ocean-vaporizing impact that does not rekindle the biosphere. With our nominal assumptions, a life-starting impact occurs 92\% or 41\% of the time when we assume a $> 4 \times 10^{20}$ kg or $> 5 \times 10^{21}$ kg impact without later ocean-vaporization is required to start life, respectively. In our worst-case scenario, which assumes that impacts can more easily vaporize the ocean, a $> 5 \times 10^{21}$ kg impact without subsequent planet sterilization occurs in 20\% of simulated impact histories (Appendix Figure \ref{fig:probabilities_of_impacts_sens}).

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/probabilities_of_impacts.pdf}
  \caption{The probability of an impact that causes favorable conditions for an origin of life on Earth. The blue solid line is the probability of an impact occurring at least once. The orange dashed line is the probability of an impact without a subsequent ocean-vaporizing collision. Probabilities are shown as a function of the minimum impact mass to produce significant prebiotic molecules (e.g. HCN). Two plausible minimum masses are the \citet{Wogan_2023} optimistic and pessimistic scenarios indicated with vertical dotted lines. The shaded region between $2 \times 10^{22}$ and $6 \times 10^{22}$ kg are Moneta-sized impacts because they deliver most all of Earth's mantle HSEs. There is a 92\% chance of an impact creating favorable origin of life conditions in the \citet{Wogan_2023} optimistic case, and a 41\% chance in the pessimistic case.}
  \label{fig:probabilities_of_impacts}
\end{figure}

@@

It is temping to extrapolate our results beyond the early Earth to exoplanets, but we must do so with caution because planet's orbiting different stars may have bombardment histories unlike the Hadean. Planets in the habitable zone of a late M-type star during the stellar main-sequence phase were interior to the habitable zone for several hundred million years during the super luminous pre-main-sequence phase \citep{Luger_2015}. \citet{Lichtenberg_2022} use N-body simulations to show that, for planets hosted by late M-dwarfs, large asteroid impacts are not useful for prebiotic chemistry because big impacts most likely occur within 100 million years of planet formation during the stellar pre-main-sequence when the planet is outside of the habitable zone. The pre-main-sequence phase is not an issue for an impact-induced origin of life on planets orbiting sun-like stars because the phase only lasts $\sim 3$ Myrs, while large asteroid impacts occur for 100s Myrs \citep{Lichtenberg_2022}. Therefore, our result that a life-starting impact was likely on the early Earth (92\% chance in an optimistic case) might be best extrapolated to habitable zone exoplanets orbiting sun-like stars.

@@

Our simulations of Earth's impact history do not always result in a bombardment favorable for an origin of life. There are some impact histories where a collision of sufficient mass to produce prebiotic molecules does not occur, or alternatively, does occur but primitive life is subsequently destroyed by an ocean-vaporizing impact that does not rekindle the biosphere. With our nominal assumptions, a life-starting impact occurs 92\% or 41\% of the time when we assume a $> 4 \times 10^{20}$ kg or $> 5 \times 10^{21}$ kg impact without later ocean-vaporization is required to start life, respectively. In our worst-case scenario, which assumes that impacts can more easily vaporize the ocean, a $> 5 \times 10^{21}$ kg impact without subsequent planet sterilization occurs in 20\% of simulated impact histories (Appendix Figure \ref{fig:probabilities_of_impacts_sens}).
